---
title: "PISA Math Analysis Round 1"
author: "Robert Glorioso"
date: "11/13/2014"
output: html_document
---

Let's look at how 15 year olds throughout the world feel about learning.  Our data comes from the Programme for International Student Assessment (PISA). Information about this organization is located at <http://www.oecd.org/pisa/pisaproducts/>. Here is an R script to get this data into R:

```{r, eval = F}
library(dplyr)
library(XML)
htmlTreeParse("http://beta.icm.edu.pl/PISAcontest/data/", useInternalNodes = T) %>% 
  xmlRoot %>% 
  xpathApply("//a", xmlGetAttr, "href") %>% 
  lapply(function(i) download.file(paste0("http://beta.icm.edu.pl/PISAcontest/data/", i), paste0(".",i)))
```

The above code will download the PISA dataset files to your current working directory.

We are most interested in the student2012 data. This is a medium sized dataset (2.7 gb). It contains 485490 observations of 635 variables. We will work with a subset of the data - students in North America:

```{r, eval = F}
load("student2012.rda")
load("student2012dict.rda")
student2012 = student2012[student2012$CNT %in% c("Florida (USA)","Connecticut (USA)", "Massachusetts (USA)", "Canada", "Mexico"),]
student2012dict = as.data.frame(sapply(student2012dict, as.character), stringsAsFactors = F) #removes factors for printing
```

Column names are encoded. A function to access the meaning of a column names is helpful:

```{r}
colDecode = function(colName){
  if(!is.numeric(colName)) colName = which(student2012dict[,1] == colName)
  if(length(colName) == 0 || colName < 1 || colName > 635) stop("Invalid Column")
  if(!is.factor(student2012[,colName])) cat(student2012dict[colName,], "\nQuantiles:\n",quantile(student2012[,colName], na.rm = T), "\n")
    else cat(student2012dict[colName,1], "\n", student2012dict[colName,2], "\nFactor Levels:\n", paste(levels(student2012[,colName]), collapse = "\n"), "\n")
}
colDecode(67)
colDecode(69)
colDecode(70)
```

We'll make a linear model with these three variables as the predictors:

```{r}
mathLM1 = lm(PV1MATH ~ ST29Q01 + ST29Q03 + ST29Q04, data = student2012)
summary(mathLM1)
```

Notice that the level "agree" is missing from the summary. That is because a linear model will absorb the first level of a categorical variable into the intercept term. We have to keep this in mind when interpreting the coefficients.

We look at the first variable "Math Interest - Enjoy Reading". It looks like there's an increase of 11 points in the score of those who answer "disagree" from those who answered "agree". We can examine a plot of this data to visualize the trend.

```{r echo = F}
library(ggplot2)
gg = ggplot(student2012, aes(x = ST29Q01, y = PV1MATH)) + geom_boxplot(colour = "blue", fill = "lightblue", notch = T, varwidth = T, outlier.size = .2, outlier.colour = "blue")
gg
```

